---
title: "Denver Broncos DSI Screening"
author: "Matthew Martinez"
date: "2023-07-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Packages
```{r Packages, warning=FALSE, message=FALSE}
library(tidyverse)
```

## Loading datasets
```{r Datasets, message=FALSE}
owners <- read_csv("owners.csv")
pets <- read_csv("pets.csv")
procedure_details <- read_csv("procedure_details.csv")
procedures <- read_csv("procedures.csv")

# Merged datasets for ease of use
ownersMergedPets <- inner_join(owners, pets, by="OwnerID")
procedure_merge_data <- inner_join(procedures, procedure_details, by = c("ProcedureType", "ProcedureSubCode"))
```

## The oldest dog in Southfield is Crocket
```{r One}
ownersMergedPets %>%
  select(City, Name.y, Kind, Age) %>%
  filter(City == "Southfield", Kind == "Dog") %>%
  .[which.max(.$Age), ]
```

## Mean number of pets per city is 1.1
```{r Two}
problem_two <- ownersMergedPets %>%
  group_by(City) %>%
  summarize(num_pets = n(), num_owners = n_distinct(OwnerID)) %>%
  mutate(mean_pets = num_pets / num_owners)
mean(problem_two$mean_pets)
```

## Who spent the most? Daniel Fay with a total of 450
```{r Three}
pet_cost <- procedure_merge_data %>%
  group_by(PetID) %>%
  summarize(total = sum(Price))
inner_join(pet_cost, ownersMergedPets, by = "PetID") %>%
  arrange(desc(total)) %>%
  slice(1) %>%
  select(Name.x, Surname)
```

## How many owners spent 20 dollars or more? 13
```{r Four} 
pet_cost %>% filter(total >= 20) %>% nrow()
```

## How many owners have at least two different kinds of pets? 4
```{r Five}
ownersMergedPets %>%
  group_by(OwnerID) %>%
  summarize(different_kinds = n_distinct(Kind)) %>%
  filter(different_kinds >= 2) %>%
  nrow()
```

## First letter of owner name matches first letter of pets name? 8
```{r Six}
ownersMergedPets %>%
  mutate(FirstLetter = substr(Name.x, 1, 1) == substr(Name.y, 1, 1)) %>%
  group_by(OwnerID) %>%
  filter(FirstLetter == TRUE) %>%
  nrow()
```

## Percentage of pets that received a vaccination? 29%
```{r Seven}
procedures %>% filter(ProcedureType == "VACCINATIONS") %>%
  nrow(.) / nrow(pets) * 100
```

## Percentage of cities with more male pets than female pets? 57.5%
```{r Eight}
cities_data <- ownersMergedPets %>%
  group_by(City, Gender) %>%
  summarize(num_pets = n()) %>%
  pivot_wider(names_from = Gender, values_from = num_pets, values_fill = 0)
cities_data %>% filter(male > female) %>%
  nrow(.) / nrow(cities_data) * 100
```

## Which city's pet sample is made up of exactly 70% dogs? Grand Rapids
```{r Nine}
ownersMergedPets %>% group_by(City, Kind) %>%
  summarize(num_pets = n()) %>%
  pivot_wider(names_from = Kind, values_from = num_pets, values_fill = 0) %>%
  mutate(total = Dog + Parrot + Cat) %>%
  mutate(percentOfDogs = (Dog / total) * 100) %>%
  filter(percentOfDogs == 70) %>%
  select(City)
```